### 2014-02-21 ThesisJP/Thesis.tex
Stil writing...

%\documentclass[11pt,a4paper]{book}
\documentclass[10pt,b5paper,twoside]{book}

% Line spacing
%\linespread{1.0}

% Include document setup script
\include{documentproperties}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TITLE PAGE
\author{Jon Petter \AA{}sen}
\title{%Integrated Ultrasound Visualization and Formation
%Real-Time Adaptive Beamforming, Visualization, and Simulation 
%Real-Time Adaptive Ultrasound Algorithms
%Accelerating adaptive ultrasound algorithms using graphics processing units 
%Accelerating adaptive ultrasound algorithms with graphics processing units 
Accelerating adaptive ultrasound algorithms by means of general-purpose programming on graphics processing units 
%Accelerating adaptive ultrasound algorithms by general purpose programming on graphics processing units 
%Accelerating ultrasound algorithms with graphics processing units 
%Accelerating ultrasound algorithms using graphics processing units 
\vspace{0.5cm}}
\department{Department of Circulation and Medical Imaging}

\usepackage{color}
\newcommand\todo[1]{{\textit{\color{red}(#1)}}}

% symbol list
\usepackage{nomencl}
\makenomenclature
\renewcommand{\nomname}{Symbols and Abbreviations}
\newcommand*{\nom}[2]{#1\nomenclature{#1}{#2}}

% custom commands
\newcommand{\mat}[1]{\mathbf{#1}}
\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\degree}{\ensuremath{^\circ}}
\newcommand{\R}{$\mat{\hat{R}}$ }

\begin{document}

\date{\today}
\pagestyle{empty} \maketitle

%Norw.  sammendrag
% Norsk sammendrag
%\Chapter*{}
\newpage%

\renewcommand{\thepage}{\roman{page}}% Roman numerals for page counter

~\\
\vspace{10.0cm}
~\\
{\bf NTNU}
\\
Norwegian University of Science and Technology\\
\\
Thesis for the degree of Philosophiae Doctor\\
\\
Faculty of Medicine\\
Department of Circulation and Medical Imaging\\
\\
\copyright\,Jon Petter \AA{}sen\\
\\
ISBN XXX-XX-XXX-XXXX-X (printed version)\\
ISBN XXX-XX-XXX-XXXX-X (electronic version)\\
ISSN XX-XX\\
\\
Doctoral theses at NTNU, 20014:XX\\
\\
Printed by NTNU-trykk

~\\
\newpage
%\thispagestyle{empty}%
~\\%
~\\%
~\\
\noindent\large 
Sammendrag p\aa{} norsk.
%\vspace{0.2cm}\\%
\\\\%
\normalsize
\todo{Skriv sammendrag p\aa{} norsk}

\vspace{0.3cm}
\noindent Jon Petter \AA{}sen\\
Institutt for sirkulasjon og bildediagnostikk, NTNU\\
Hovedveileder: Sverre Holm\\ 
Biveiledere: Stein-Inge Rabben, Erik Steen and Hans Torp\\
Finansieringskilde: MI-Lab

\vspace{0.2cm}
\noindent \emph{Ovennevnte avhandling er funnet verdig til \aa{} forsvares offentlig for graden Philosophiae Doctor (PhD) i medisinsk teknologi. Disputas finner sted ... .}

% ENGLISH ABSTRACT / SUMMARY
\CustChapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
%\chapter*{Abstract}
%\thispagestyle{empty}%
\todo{Write abstract ...}

%\setcounter{page}{3} %I don't want the norwegian summary to be counted in. (This will change page numbering since the delivered version)

% PREFACE AND ACKNOWLEDGEMENTS
\CustChapter*{Preface}%
\addcontentsline{toc}{chapter}{Preface}
%\thispagestyle{empty}%
The present thesis is submitted in partial fulfillment of the requirements for the degree of \textit{Philosophiae Doctor} (Ph.D.) at the Faculty of Medicine at the Norwegian University of Science and Technology (NTNU). The research was funded by \textit{Medical Imaging Laboratory} (MI-Lab), and was carried out at the Department of Circulation and Medical Imaging. The main supervisor has been Professor Sverre Holm  in the Department of Informatics, University of Oslo, and Professor II at MI-Lab, NTNU. Co-supervisors have been PhD Stein Inge Rabben, PhD Erik Steen and Professor Hans Torp.

\subsubsection{Acknowledgments}
\todo{Write acknowledgments ...}

\CustChapter*{List of publications}
\addcontentsline{toc}{chapter}{List of publications}

This thesis is based on the following five papers, referred to in the text by their Roman numerals (I-V): 
\begin{enumerate}[I]
\renewcommand\labelenumi{\bfseries\theenumi}
	\item J.\:I.\:Buskenes, \textbf{J.\:P.\:\AA{}sen}, C.-I.\:C.\:Nilsen and A.\:Austeng, "An optimised GPU implementation of the MVDR beamformer for active sonar imaging", {\it  IEEE Transactions on Oceanic Engineering, accepted for publication}.
	\item \textbf{J.\:P.\:\AA{}sen}, J.\:I.\:Buskenes, C.-I.\:C.\:Nilsen, A.\:Austeng and S.\:Holm, "Implementing Capon beamforming on a GPU for real-time cardiac ultrasound imaging", {\it IEEE Transactions on Ultrasonics, Ferroelectrics, and Frequency Control, vol. 61, no. 1, pp. 76-85, Jan 2014}.
 	\item \textbf{J.\:P.\:\AA{}sen}, A.\:Austeng and S.\:Holm, "Capon beamforming and moving objects - An analysis of lateral shift-invariance", {\it IEEE Transactions on Ultrasonics, Ferroelectrics, and Frequency Control, accepted for publication}.
	\item \textbf{J.\:P.\:\AA{}sen}, E.\:Steen, G.\:Kiss, A.\:Thorstensen and S.\:I.\:Rabben, "Adaptive volume rendering of cardiac 3D ultrasound images - utilizing blood pool statistics", {\it Proc. SPIE Medical Imaging 2012, vol. 8320, pp. 832008}.
	\item \textbf{J.\:P.\:\AA{}sen} and S.\:Holm, "Huygens on speed: Interactive simulation of ultrasound pressure fields", {\it Proc. IEEE Ultrasonics Symposium 2012, pp. 1643-1646}.
\end{enumerate}

\CustChapter*{Related publications}
\addcontentsline{toc}{chapter}{Related publications}
The following papers are related to the included papers, and are referred to in the text by their Roman numerals:

\begin{enumerate}[VI]
\setcounter{enumi}{5}
\renewcommand\labelenumi{\bfseries\theenumi}

	\item \textbf{J.\:P.\:\AA{}sen}, J.\:I.\:Buskenes, C.-I.\:C.\:Nilsen, A.\:Austeng and S.\:Holm, "Implementing Capon Beamforming on the GPU for Real-Time Cardiac Ultrasound Imaging", {\it Proc. IEEE Ultrasonics Symposium 2012}, pp. 2133-2136.

	\item J.\:I.\:Buskenes, \textbf{J.\:P.\:\AA{}sen}, C.-I.\:C.\:Nilsen and A.\:Austeng, "Adapting the Minimum variance beamformer to a graphics processing unit for active sonar imaging systems", {\it Proc. Meetings on Acoustics 2013} vol. 13.

	\item G.\:Kiss, E.\:Steen, \textbf{J.\:P.\:\AA{}sen} and H.\:Torp, "GPU volume rendering in 3D echocardiography: real-time pre-processing and ray-casting", {\it Proc. IEEE Ultrasonics Symposium 2010}, pp. 193-196.

	%\item G. Kiss, \textbf{J. P. \AA{}sen}, J. Bogaert, B. H. Amundsen, P. Claus, J. R. M. D'hooge and H. Torp, "Multi-modal cardiac image fusion and visualization on the GPU", {\it Proc. IEEE Ultrasonics Symposium 2011}.

	\item E.\:S.\:Br\o{}nstad, \textbf{J.\:P.\:\AA{}sen}, H.\:Torp and G.\:Kiss, "Visibility driven visualization of 3D cardiac ultrasound data on the GPU", {\it Proc. IEEE Ultrasonics Symposium 2012}, pp. 2651-2654.
\end{enumerate}

\newpage\thispagestyle{empty}
\pagestyle{fancy}

%

\cleardoublepage
\addcontentsline{toc}{chapter}{Contents}
\renewcommand*\contentsname{Contents}
\tableofcontents

\cleardoublepage
\addcontentsline{toc}{chapter}{Symbols and Abbreviations}
\printnomenclature

\cleardoublepage
\setcounter{page}{1}
\renewcommand{\thepage}{\arabic{page}}

%\renewcommand*\thesection{\arabic{section}}

%%%% CHAPTER 2 %%%%
\begin{bibunit}[ieeetr]
\CustChapter{Background}
\runningtitle{Background}
\input{./Background/background}
\newpage\addcontentsline{toc}{section}{References}
\putbib[./Background/background]
\end{bibunit}

%\setcounter{section}{0}
%\setcounter{chapter}{1}

%%%% CHAPTER 1 %%%%
\CustChapter{Introduction} 
\runningtitle{Introduction}
\input{./Introduction/introduction}

\setcounter{chapter}{0}

%%%% Paper 1 %%%%
\begin{bibunit}[ieeetr]
\CustChapterPaper{An Optimised GPU Implementation of the MVDR Beamformer for Active Sonar Imaging}
\runningtitle{GPU MVDR Beamformer for Active Sonar Imaging}
\authors{
	Jo Inge Buskenes$^{1}$, 
    	\textbf{Jon Petter \AA{}sen}$^{2}$, 
    	Carl-Inge C. Nilsen$^{1}$ and
    	Andreas Austeng$^{1}$
} 
{
	$^{1}$Department of Informatics, University of Oslo, Oslo, Norway\\
    	$^{2}$Medical Imaging Lab (MI-Lab), Norwegian University of Science and Technology, Trondheim, Norway
}
\noindent \textit{IEEE Transactions on Oceanic Engineering, accepted for publication}.
%\newpage\input{./CaponSonarGPU/CaponSonarGPU}
%\newpage\addcontentsline{toc}{section}{References}
%\putbib[./CaponSonarGPU/bibCaponSonarGPU]
\end{bibunit}

%%%% Paper 2 %%%%
\begin{bibunit}[ieeetr]
\CustChapterPaper{Implementing Capon Beamforming on a GPU for Real-Time Cardiac Ultrasound Imaging}
\runningtitle{GPU Capon Beamformer for Cardiac Ultrasound Imaging}
\authors{
	\textbf{Jon Petter \AA{}sen}$^{1}$,
    	Jo Inge Buskenes$^{2}$,  
    	Carl-Inge C. Nilsen$^{2}$,
    	Andreas Austeng$^{2}$ and
    	Sverre Holm$^{1,2}$
} 
{
	$^{1}$Medical Imaging Lab (MI-Lab), Norwegian University of Science and Technology, Trondheim, Norway\\
    	$^{2}$Department of Informatics, University of Oslo, Oslo, Norway
}
\noindent \textit{IEEE Transactions on Ultrasonics, Ferroelectrics, and Frequency Control, vol. 61, pp. 76-85, Jan 2014}.
%\input{./CaponCardiacGPU/CaponCardiacGPU}
%\newpage\addcontentsline{toc}{section}{References}
%\putbib[./CaponCardiacGPU/bibCaponCardiacGPU]
\end{bibunit}

%%%% Paper 3 %%%%
\begin{bibunit}[ieeetr]
\CustChapterPaper{Capon Beamforming and Moving Objects - An Analysis of Lateral Shift-Invariance}
\runningtitle{Capon Beamforming and Moving Objects}
\authors{
	\textbf{Jon Petter \AA{}sen}$^{1}$, 
    	Andreas Austeng$^{2}$ and
    	Sverre Holm$^{1,2}$
} 
{
	$^{1}$Medical Imaging Lab (MI-Lab), Norwegian University of Science and Technology, Trondheim, Norway\\
    	$^{2}$Department of Informatics, University of Oslo, Oslo, Norway
}
\noindent \textit{IEEE Transactions on Ultrasonics, Ferroelectrics, and Frequency Control, accepted for publication}.
%\input{./CaponMovingObjects/CaponMovingObjects}
%\newpage\addcontentsline{toc}{section}{References}
%\putbib[./CaponMovingObjects/bibCaponMovingObjects]
\end{bibunit}

%%%% Paper 4 %%%%
\begin{bibunit}[ieeetr]
\CustChapterPaper{Adaptive Volume Rendering of Cardiac 3D Ultrasound Images - Utilizing Blood Pool Statistics}
\runningtitle{Adaptive Volume Rendering of Cardiac Ultrasound}
\authors{
	\textbf{Jon Petter \AA{}sen}$^{1}$,
	Erik Steen$^2$,
	Gabriel Kiss$^{3,4}$,
	Anders Thorstensen$^{1,5}$,
	Stein Inge Rabben$^2$
}
{
$^{1}$Medical Imaging Lab (MI-Lab), Norwegian University of Science and Technology, Trondheim, Norway\\
$^{2}$ GE Vingmed Ultrasound, Horten, Norway\\
$^{3}$ Norwegian University of Science and Technology, ISB - Department of Circulation and Medical Imaging, Trondheim, Norway\\
$^{4}$ St. Olavs Hospital, Trondheim, Norway\\
$^{5}$ Department of Cardiology, St. Olavs Hospital, Trondheim, Norway
}
\noindent \textit{Proc. SPIE Medical Imaging 2012, vol. 8320, pp. 832008.}
%\input{./AdaptiveVolumeRendering/AdaptiveVolumeRendering}
%\newpage\addcontentsline{toc}{section}{References}
%\putbib[./AdaptiveVolumeRendering/bibAdaptiveVolumeRendering]
\end{bibunit}

%%%% Paper 5 %%%%
\begin{bibunit}[ieeetr]
\CustChapterPaper{Huygens on Speed: Interactive Simulation of Ultrasound Pressure Fields}
\runningtitle{Huygens on speed}
\authors{\textbf{Jon Petter \AA{}sen}$^1$ and Sverre Holm$^{1,2}$}
{
	$^{1}$Medical Imaging Lab (MI-Lab), Norwegian University of Science and Technology, Trondheim, Norway\\
    	$^{2}$Department of Informatics, University of Oslo, Oslo, Norway
}
\noindent \textit{Proc. IEEE Ultrasonics Symposium 2012, pp. 1643-1646}.
%\input{./Hos/Hos}
%\newpage\addcontentsline{toc}{section}{References}
%\putbib[./Hos/bibHos]
\end{bibunit}

%\appendix

\end{document}