### 2014-04-13 ThesisJP/Thesis.tex

%\documentclass[11pt,a4paper]{book}
\documentclass[10pt,b5paper,twoside]{book}

% Line spacing
%\linespread{1.0}

% Include document setup script
\include{documentproperties}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TITLE PAGE
\author{Jon Petter Helgesen \AA{}sen}
\title{%Integrated Ultrasound Visualization and Formation
%Real-Time Adaptive Beamforming, Visualization, and Simulation 
%Real-Time Adaptive Ultrasound Algorithms
%Accelerating adaptive ultrasound algorithms using graphics processing units 
%Accelerating adaptive ultrasound algorithms with graphics processing units 
Accelerating adaptive ultrasound imaging algorithms by means of general-purpose computing on graphics processing units 
%Accelerating adaptive ultrasound algorithms by general purpose programming on graphics processing units 
%Accelerating ultrasound algorithms with graphics processing units 
%Accelerating ultrasound algorithms using graphics processing units 
%Pinner og romber fra det yttre rom
\vspace{0.5cm}}
\department{Department of Circulation and Medical Imaging}

\usepackage{color}
\newcommand\todo[1]{{\textit{\color{red}(#1)}}}

% from Paper 3
\usepackage{epstopdf}

% from spie.tex (Paper 4)
\usepackage{array}

% from hos (Paper 5)
\usepackage{url}
%\usepackage{epsfig}

% symbol list
\usepackage{nomencl}
\makenomenclature
\renewcommand{\nomname}{Symbols and Abbreviations}
\newcommand*{\nom}[2]{#1\nomenclature{#1}{#2}}
\newcommand*{\nomi}[2]{\nomenclature{#1}{#2}}

% custom commands
\newcommand{\mat}[1]{\mathbf{#1}}
\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\degree}{\ensuremath{^\circ}}
\newcommand{\R}{$\mat{\hat{R}}$ }

\begin{document}

\date{\today}
\pagestyle{empty} \maketitle

%Norw.  sammendrag
% Norsk sammendrag
%\Chapter*{}
\newpage%

\renewcommand{\thepage}{\roman{page}}% Roman numerals for page counter

~\\
\vspace{10.0cm}
~\\
{\bf NTNU}
\\
Norwegian University of Science and Technology\\
\\
Thesis for the degree of Philosophiae Doctor\\
\\
Faculty of Medicine\\
Department of Circulation and Medical Imaging\\
\\
\copyright\,Jon Petter \AA{}sen\\
\\
ISBN XXX-XX-XXX-XXXX-X (printed version)\\
ISBN XXX-XX-XXX-XXXX-X (electronic version)\\
ISSN XX-XX\\
\\
Doctoral theses at NTNU, 20014:XX\\
\\
Printed by NTNU-trykk

~\\
\newpage
%\thispagestyle{empty}%
~\\%
~\\%
~\\
\noindent\large 
Sammendrag p\aa{} norsk.
%\vspace{0.2cm}\\%
\\\\%
\normalsize
\todo{Skriv sammendrag p\aa{} norsk}

\vspace{0.3cm}
\noindent Jon Petter Helgesen \AA{}sen\\
Institutt for sirkulasjon og bildediagnostikk, NTNU\\
Hovedveileder: Sverre Holm\\ 
Biveiledere: Stein-Inge Rabben, Erik Steen and Hans Torp\\
Finansieringskilde: MI-Lab

\vspace{0.2cm}
\noindent \emph{Ovennevnte avhandling er funnet verdig til \aa{} forsvares offentlig for graden Philosophiae Doctor (PhD) i medisinsk teknologi. Disputas finner sted ... .}

% ENGLISH ABSTRACT / SUMMARY
\CustChapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
%\chapter*{Abstract}
%\thispagestyle{empty}%
\todo{Write abstract ...}

%\setcounter{page}{3} %I don't want the norwegian summary to be counted in. (This will change page numbering since the delivered version)

% PREFACE AND ACKNOWLEDGEMENTS
\CustChapter*{Preface}%
\addcontentsline{toc}{chapter}{Preface}
%\thispagestyle{empty}%
The present thesis is submitted in partial fulfillment of the requirements for the degree of \textit{Philosophiae Doctor} (Ph.D.) at the Faculty of Medicine at the Norwegian University of Science and Technology (NTNU). The research was funded by \textit{Medical Imaging Laboratory} (MI-Lab), and was carried out at the Department of Circulation and Medical Imaging. The main supervisor has been Professor Sverre Holm in the Department of Informatics, University of Oslo (UiO), and Professor II at MI-Lab, NTNU. Co-supervisors have been Ph.D. Stein Inge Rabben, Ph.D. Erik Steen, both at GE Vingmed Ultrasound AS, and Professor Hans Torp in the Department of Circulation and Medical Imaging, NTNU.

\subsubsection{Acknowledgments}
First of all I would like to thank my main supervisor Sverre Holm and the rest of my team of supervisors for all help and guidance through the Ph.D project. Andreas Austeng also deserves big thank you. In the second part of my Ph.D. project you have been my supervisor more than anyone. Thank you to all my co-authors for their contributions. I also want to thank the Faculty of Medicine, MI-Lab and the Department of Circulation and Medical Imaging, NTNU for believing in me and making it possible to conduct this work.

During these years I have had the great pleasure of working with and getting to know several research communities and companies. I therefore would like to thank all colleagues at the Department of Circulation and Medical Imaging, NTNU, the Department of Informatics, UiO, GE Vingmed Ultrasound AS, and Squarehead Technology AS for providing great working environments. A special thanks goes to all my fellow Ph.D. students at MI-Lab. I will never forget our conference trips. A special thanks also goes to Jo Inge Buskenes at "office" 4470 for making every working hour so much fun.

Thank you to Ines Hafizovic and Carl-Inge Colombo Nilsen for extensive help with reviewing this thesis.

I would like to thank my family for all support. But most of all, I want to thank Beate for all her love, patient and support. Without it I would never have manage to complete this theses.
\null
\vfill
\hfill Dedicated to Sofie Mathilde \AA{}sen
\CustChapter*{List of publications}
\addcontentsline{toc}{chapter}{List of publications}

This thesis is based on the following five papers, referred to in the text by their Roman numerals (I-V): 
\begin{enumerate}[I]
\renewcommand\labelenumi{\bfseries\theenumi}
	\item J.\:I.\:Buskenes, \textbf{J.\:P.\:\AA{}sen}, C.-I.\:C.\:Nilsen and A.\:Austeng, "An optimised GPU implementation of the MVDR beamformer for active sonar imaging", {\it  IEEE Transactions on Oceanic Engineering, accepted for publication}.
	\item \textbf{J.\:P.\:\AA{}sen}, J.\:I.\:Buskenes, C.-I.\:C.\:Nilsen, A.\:Austeng and S.\:Holm, "Implementing Capon beamforming on a GPU for real-time cardiac ultrasound imaging", {\it IEEE Transactions on Ultrasonics, Ferroelectrics, and Frequency Control, vol. 61, no. 1, pp. 76-85, Jan 2014}.
 	\item \textbf{J.\:P.\:\AA{}sen}, A.\:Austeng and S.\:Holm, "Capon beamforming and moving objects - An analysis of lateral shift-invariance", {\it IEEE Transactions on Ultrasonics, Ferroelectrics, and Frequency Control, accepted for publication}.
	\item \textbf{J.\:P.\:\AA{}sen}, E.\:Steen, G.\:Kiss, A.\:Thorstensen and S.\:I.\:Rabben, "Adaptive volume rendering of cardiac 3D ultrasound images - utilizing blood pool statistics", {\it Proc. SPIE Medical Imaging 2012, vol. 8320, pp. 832008}.
	\item \textbf{J.\:P.\:\AA{}sen} and S.\:Holm, "Huygens on speed: Interactive simulation of ultrasound pressure fields", {\it Proc. IEEE Ultrasonics Symposium 2012, pp. 1643-1646}.
\end{enumerate}

\CustChapter*{Related publications}
\addcontentsline{toc}{chapter}{Related publications}
The following papers are related to the included papers, and are referred to in the text by their Roman numerals:

\begin{enumerate}[VI]
\setcounter{enumi}{5}
\renewcommand\labelenumi{\bfseries\theenumi}

	\item \textbf{J.\:P.\:\AA{}sen}, J.\:I.\:Buskenes, C.-I.\:C.\:Nilsen, A.\:Austeng and S.\:Holm, "Implementing Capon Beamforming on the GPU for Real-Time Cardiac Ultrasound Imaging", {\it Proc. IEEE Ultrasonics Symposium 2012, pp. 2133-2136}.

	\item J.\:I.\:Buskenes, \textbf{J.\:P.\:\AA{}sen}, C.-I.\:C.\:Nilsen and A.\:Austeng, "Adapting the Minimum variance beamformer to a graphics processing unit for active sonar imaging systems", {\it Proc. Meetings on Acoustics 2013, vol. 13}. \textit{Invited talk}.

	\item G.\:Kiss, E.\:Steen, \textbf{J.\:P.\:\AA{}sen} and H.\:Torp, "GPU volume rendering in 3D echocardiography: real-time pre-processing and ray-casting", {\it Proc. IEEE Ultrasonics Symposium 2010, pp. 193-196}.

	%\item G. Kiss, \textbf{J. P. \AA{}sen}, J. Bogaert, B. H. Amundsen, P. Claus, J. R. M. D'hooge and H. Torp, "Multi-modal cardiac image fusion and visualization on the GPU", {\it Proc. IEEE Ultrasonics Symposium 2011}.

	\item E.\:S.\:Br\o{}nstad, \textbf{J.\:P.\:\AA{}sen}, H.\:Torp and G.\:Kiss, "Visibility driven visualization of 3D cardiac ultrasound data on the GPU", {\it Proc. IEEE Ultrasonics Symposium 2012, pp. 2651-2654}.
\end{enumerate}

\newpage\thispagestyle{empty}
\pagestyle{fancy}

%

\cleardoublepage
\addcontentsline{toc}{chapter}{Contents}
\renewcommand*\contentsname{Contents}
\tableofcontents

\cleardoublepage
\addcontentsline{toc}{chapter}{Symbols and Abbreviations}
\printnomenclature

\cleardoublepage
\setcounter{page}{1}
\renewcommand{\thepage}{\arabic{page}}

%\renewcommand*\thesection{\arabic{section}}

%%%% CHAPTER 2 %%%%
\begin{bibunit}[IEEEtran]
\CustChapter{Background}
\runningtitle{Background}
\input{./Background/background}
\newpage\addcontentsline{toc}{section}{References}
\putbib[./Background/background]
\end{bibunit}

%\setcounter{section}{0}
%\setcounter{chapter}{1}

%%%% CHAPTER 1 %%%%
\CustChapter{Introduction} 
\runningtitle{Introduction}
\input{./Introduction/introduction}

\setcounter{chapter}{0}

%%%% Paper 1 %%%%
\begin{bibunit}[IEEEtran]
\CustChapterPaper{An Optimised GPU Implementation of the MVDR Beamformer for Active Sonar Imaging}
\runningtitle{GPU MVDR Beamformer for Active Sonar Imaging}
\authors{
	Jo Inge Buskenes$^{1}$, 
    	\textbf{Jon Petter \AA{}sen}$^{2}$, 
    	Carl-Inge C. Nilsen$^{1}$ and
    	Andreas Austeng$^{1}$
} 
{
	$^{1}$Department of Informatics, University of Oslo, Oslo, Norway\\
    	$^{2}$Medical Imaging Lab (MI-Lab), Norwegian University of Science and Technology, Trondheim, Norway
}
\noindent \textit{IEEE Transactions on Oceanic Engineering, accepted for publication}.
%\newpage\input{./CaponSonarGPU/CaponSonarGPU}
%\newpage\addcontentsline{toc}{section}{References}
%\putbib[./CaponSonarGPU/bibCaponSonarGPU]
\end{bibunit}

%%%% Paper 2 %%%%
\begin{bibunit}[IEEEtran]
\CustChapterPaper{Implementing Capon Beamforming on a GPU for Real-Time Cardiac Ultrasound Imaging}
\runningtitle{GPU Capon Beamformer for Cardiac Ultrasound Imaging}
\authors{
	\textbf{Jon Petter \AA{}sen}$^{1}$,
    	Jo Inge Buskenes$^{2}$,  
    	Carl-Inge C. Nilsen$^{2}$,
    	Andreas Austeng$^{2}$ and
    	Sverre Holm$^{1,2}$
} 
{
	$^{1}$Medical Imaging Lab (MI-Lab), Norwegian University of Science and Technology, Trondheim, Norway\\
    	$^{2}$Department of Informatics, University of Oslo, Oslo, Norway
}
\noindent \textit{IEEE Transactions on Ultrasonics, Ferroelectrics, and Frequency Control, vol. 61, pp. 76-85, Jan 2014}.
\newpage
{\input{./Paper2/cos}}
\newpage
\addcontentsline{toc}{section}{References}
\putbib[./Paper2/cosbib]
\end{bibunit}

%%%% Paper 3 %%%%
\begin{bibunit}[IEEEtran]
\CustChapterPaper{Capon Beamforming and Moving Objects - An Analysis of Lateral Shift-Invariance}
\runningtitle{Capon Beamforming and Moving Objects}
\authors{
	\textbf{Jon Petter \AA{}sen}$^{1}$, 
    	Andreas Austeng$^{2}$ and
    	Sverre Holm$^{1,2}$
} 
{
	$^{1}$Medical Imaging Lab (MI-Lab), Norwegian University of Science and Technology, Trondheim, Norway\\
    	$^{2}$Department of Informatics, University of Oslo, Oslo, Norway
}
\noindent \textit{IEEE Transactions on Ultrasonics, Ferroelectrics, and Frequency Control, accepted for publication}.
\newpage\input{./Paper3/moving_obj}
\newpage\addcontentsline{toc}{section}{References}
\putbib[./Paper3/moving_objbib]
\end{bibunit}

%%%% Paper 4 %%%%
\begin{bibunit}[IEEEtran]
\CustChapterPaper{Adaptive Volume Rendering of Cardiac 3D Ultrasound Images - Utilizing Blood Pool Statistics}
\runningtitle{Adaptive Volume Rendering of Cardiac Ultrasound}
\authors{
	\textbf{Jon Petter \AA{}sen}$^{1}$,
	Erik Steen$^2$,
	Gabriel Kiss$^{3,4}$,
	Anders Thorstensen$^{1,5}$,
	Stein Inge Rabben$^2$
}
{
$^{1}$Medical Imaging Lab (MI-Lab), Norwegian University of Science and Technology, Trondheim, Norway\\
$^{2}$ GE Vingmed Ultrasound, Horten, Norway\\
$^{3}$ Norwegian University of Science and Technology, ISB - Department of Circulation and Medical Imaging, Trondheim, Norway\\
$^{4}$ St. Olavs Hospital, Trondheim, Norway\\
$^{5}$ Department of Cardiology, St. Olavs Hospital, Trondheim, Norway
}
\noindent \textit{Proc. SPIE Medical Imaging 2012, vol. 8320, pp. 832008.}
\newpage\input{./Paper4/spie}
\newpage\addcontentsline{toc}{section}{References}
\putbib[./Paper4/spiebib]
\end{bibunit}

%%%% Paper 5 %%%%
\begin{bibunit}[IEEEtran]
\CustChapterPaper{Huygens on Speed: Interactive Simulation of Ultrasound Pressure Fields}
\runningtitle{Huygens on speed}
\authors{\textbf{Jon Petter \AA{}sen}$^1$ and Sverre Holm$^{1,2}$}
{
	$^{1}$Medical Imaging Lab (MI-Lab), Norwegian University of Science and Technology, Trondheim, Norway\\
    	$^{2}$Department of Informatics, University of Oslo, Oslo, Norway
}
\noindent \textit{Proc. IEEE Ultrasonics Symposium 2012, pp. 1643-1646}.
\newpage\input{./Paper5/hos}
\newpage\addcontentsline{toc}{section}{References}
\putbib[./Paper5/hosbib]
\end{bibunit}

%\appendix

\end{document}