### 2014-04-23 HowtoHowtoCapon/bio/andreas_austeng.tex
JOE Final submit

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bio/andreas.eps}}]{Andreas Austeng}
was born in Oslo, Norway, in 1970. He received the M.Sc. degree in physics in 1996 and the Ph.D. degree in computer science in 2001, both from the University of Oslo. Since 2001, he has been working at the Department of Informatics, University of Oslo, first as a postdoctoral research fellow and currently as an associate professor. His research interests include signal and array processing for acoustical imaging.
\end{IEEEbiography}