### 2016-02-16 LCAforActiveSonarImaging/lca_formula.tex
Working on the review

\newpage
%
The angular frequency response of a window steered to $\phi'$ ($\lambda$ disappears if we express element distance $d$ relative to the wavelength):
%
\begin{align*}
W_{M,d,\beta,\phi'}(\phi) &= \sum\limits_{m=0}^{M-1} w_{m,M,d,\beta,\phi'}\, e^{j\frac{2\pi}{\lambda}md\sin\phi} \\
&= \sum\limits_{m=0}^{M-1} e^{-j\frac{2\pi}{\lambda}md\sin\phi'}\,f_{m,M}(\beta)\,e^{j\frac{2\pi}{\lambda}md\sin\phi} \\
&= \sum\limits_{m=0}^{M-1} f_{m,M}(\beta)\,e^{j\frac{2\pi}{\lambda}md\big[\sin\phi'-\sin\phi\big]}
\end{align*}
%
Or, expressed as a function of wavenumber $k_x = \|\vec k\|\sin\phi = \frac{2\pi}{\lambda}\sin\phi$:
%
\begin{align*}
W_{M,d,\beta,k_x'}(k_x) &= \sum\limits_{m=0}^{M-1} w_{m,M,d,\beta,\phi'}\, e^{jk_xmd} \\
&= \sum\limits_{m=0}^{M-1} e^{-jk_x'(\phi) md}\,f_{m,M}(\beta)\,e^{jk_xmd} \\
&= \sum\limits_{m=0}^{M-1} f_{m,M}(\beta)\,e^{j\big[k_x'(\phi)-k_x\big]md}
\end{align*}
%
What I want to know is what steering $k_x'$ that causes the amplitude to be -3dB at some predetermined location $k_x^0$, under the constraint that weights always sum to 1 (for unit gain at $k_x=0$):
%
\begin{align*}
\Big|W_{M,d,\beta,k'}(k_x^0)\Big| &= \frac{1}{\sqrt{2}} \quad\quad\text{subject to:}\quad\quad \sum\limits_{m=0}^{M-1} w_{m,M,d,\beta,\phi'} = 1.
\end{align*}
%
Is this possible to solve somehow? I don't mind if it needs to be a numeric solution or an approximation. I need it in my article and some kind of formula would be better than the algorithm I have now that performs a search to get the answer. I was wondering if the equation can be separated into a term/factor that independently depends on $M$, $d$ and $\beta$ somehow, then numerically solve each and create a compounded equation/approximation of the results, but my math skills are struggling :P

\newpage