### 2018-02-06 LCAforActiveSonarImaging/diagram/overview.tex

\documentclass[tikz]{standalone}

\usepackage{pgf}
% % \usepackage{pgf-blur}
\usepackage{tikz}
\usetikzlibrary{arrows,automata,graphs,
		fadings,mindmap,positioning,
		shadows,shadows.blur,
	        shapes} 
% \usegdlibrary{layered}

\pagestyle{empty}

\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
    text width=4.5em, text badly centered, node distance=3cm, inner sep=0pt]

\tikzstyle{section} = [
   draw,
   ellipse,
   fill=gray!50!gray,
%    drop shadow,
   blur shadow={shadow blur steps=8},
   node distance=3cm,
   minimum height=2em,
   shade, shading=axis, top color=black!10!white, bottom color=black!25!white %shading angle=45
   ]

\tikzstyle{goal} = [
%    draw,
%    rounded rectangle,
%    fill=white,
%    drop shadow,
%    blur shadow={shadow blur steps=8},
   node distance=0.5cm,
   minimum height=0.2cm,
   minimum height=0.2cm,
   ]

\tikzstyle{structure} = [
   draw, ellipse,
   fill=white!20,
   drop shadow,
   node distance=3cm,
   minimum height=2em]

\tikzset{state/.style={
	rectangle,
	rounded corners,
	draw=black, very thick,
	minimum height=2em,
	inner sep=2pt,
	text centered,
	},
}

\begin{document}
\begin{tikzpicture}[->,>=stealth']

\node[section](abstract){ \bf Abstract };

\node[section, below of=abstract](introduction){ \bf Introduction };

\node[section, below of=introduction](methods){ \bf Methods };

\node[section, below of=methods](results){ \bf Results };

\node[section, below of=results](discussion){ \bf Discussion };

\node[section, below of=discussion](conclusion){ \bf Conclusion };

\node[goal, left of=introduction](intro_goals){ \bf Goals };
\node[goal, left of=intro_goals](intro_goals_why){ \bf Why };
\node[goal, below of=intro_goals_why](intro_goals_what){ \bf What };

\path
(abstract)        edge  (introduction)
(introduction)    edge  (methods)
   (introduction) edge  (goals)
(methods)         edge  (results)
(results)         edge  (discussion)
(discussion)      edge  (conclusion);

\end{tikzpicture}
\end{document}