### 2018-02-06 Simulator/bio/jo_inge_buskenes_new.tex

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bio/jo_inge.eps}}]{Jo Inge Buskenes}
received the B.Sc. degree in electrical engineering from Gj\o{}vik College University, Norway, in 2007, and the M.Sc. degree in instrumentation for particle physics from the University of Oslo, Norway, in 2010. He is currently pursuing the Ph.D. degree in acoustic image reconstruction and high performance computing at the University of Oslo.

His industry experience includes development of digital electronics at the European Organization for Nuclear Research (CERN), Geneva, Switzerland (2007-2008). He has lectured in digital signal processing at the Gj\o{}vik College University (2009), and at the University of Oslo (2010-2013). Current affiliation is with The Norwegian Defence Research Establishment, Kjeller, Norway, for which he is developing radar systems (2015-), and formerly sonar systems (2009, 2013).

His research interests include radar and sonar technology, adaptive image reconstruction, high performance computing, intelligent detector design and open source software.
\end{IEEEbiography}