### 2018-02-06 Simulator/bio/roy_edgar_hansen.tex

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bio/roy_color.eps}}]{Roy Edgar Hansen} (M'07) received the M.Sc. and Ph.D. degrees in physics from the University of Troms\o, Troms\o, Norway, in 1992 and 1999, respectively.

From 1992 to 2000, he was with the Norwegian research company TRIAD, working on multistatic sonar, multistatic radar, synthetic aperture radar (SAR), and underwater communications. Since 2000, he has been with the Norwegian Defence Research Establishment (FFI), Kjeller, Norway. He is currently Principal Scientist and team leader for the autonomous underwater vehicle development and the synthetic aperture sonar development at FFI. He is also Adjunct Associated Professor at the University of Oslo, Oslo, Norway.
\end{IEEEbiography}