### 2018-02-06 Simulator/bio/roy_edgar_hansen_new.tex

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bio/roy_color.eps}}]{Roy Edgar Hansen} (M'07) received the M.Sc. and Ph.D. degrees in physics from the University of Troms\o, Troms\o, Norway, in 1992 and 1999, respectively.
The PhD thesis is named Measurements in the Mixed Layer by a Bistatic Multi-CW Doppler Sonar. 
From 1992 to 2000, he was with the Norwegian research company TRIAD, working on multistatic sonar, multistatic radar, synthetic aperture radar (SAR), and underwater communications. Since 2000, he has been with the Norwegian Defence Research Establishment (FFI), Kjeller, Norway. 
In the period 2009-2015, he was research manager for the autonomous underwater vehicle development at FFI, including the fields underwater navigation, battery and fuel-cell technologies, decisional autonomy, and synthetic aperture sonar. 
He is currently principal scientist at FFI. 
He is also adjunct associated professor at Department of informatics at University in Oslo, Norway.
His research interests include synthetic aperture sonar and radar, ultrasound imaging, and array signal processing.
\end{IEEEbiography}