### 2018-02-16 ThesisJo/literature_review.tex

\input{variables}\ifMonolithic\else\input{header}\input{subheader}\fi
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\chapter{Litterature Review}

\section{Simulators}

\begin{itemize}
\item SST - Sonar Simulation Toolkit. \cite{Goddard}
\end{itemize}

General Modeling tools
\begin{itemize}
\item CASS - Comprehensive Acoustic System Simulation. \cite{Goddard}
\end{itemize}

GSM - Generic Sonar Model.

FOCUS simulator.

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\ifMonolithic\else\input{subfooter}\fi