### 2018-03-05 ThesisJo/3_discussion.tex

%&header
% !TeX root = 3_discussion
\documentclass[thesis.tex]{subfiles}
\input{subheader}
\begin{document}

\chapter{Discussion and Outlook}

\begin{itemize}
\item Improving imaging systems by making them smarter.
\item Big data - more processing power.
\item Processing speed
\begin{itemize}
	\item Choose the best tool (algorithm) for the job
	\item More complex not always better, but slower
	\item Extracting the essential information is key 
	\item Exploit hardware resources available
\end{itemize}
\item Faster IS smarter
 \begin{itemize}
 	\item Short feedback loops allow real-time adaptivity
 	\item More data contains more information
 	\item Faster learning
 \end{itemize}
\item Smarter IS better
 \begin{itemize}
	\item Better image quality
	\item Improved autonomy
\end{itemize}
\end{itemize}

\begin{itemize}
\item Improving imaging systems by making them smarter.
\item Big data - more processing power.
\item IV - Simulator
\begin{itemize}
	\item Incredibly fast template matching possible.
	\item Possibly a powerful tool to pass prior knowledge to deep learning
	\item Extracting the essential information is key 
	\item Exploit hardware resources available
\end{itemize}
\item Faster IS smarter
 \begin{itemize}
 	\item Short feedback loops allow real-time adaptivity
 	\item More data contains more information
 	\item Faster learning
 \end{itemize}
\item Smarter IS better
 \begin{itemize}
	\item Better image quality
	\item Improved autonomy
\end{itemize}
\end{itemize}

\end{document}