### 2018-03-05 ThesisJo/literature_review.tex

\input{variables}\ifMonolithic\else\input{header}\input{subheader}\fi
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\chapter{Litterature Review}

\section{Simulators}

\begin{itemize}
\item SST - Sonar Simulation Toolkit. \cite{Goddard}
\end{itemize}

General Modeling tools
\begin{itemize}
\item CASS - Comprehensive Acoustic System Simulation. \cite{Goddard}
\end{itemize}

GSM - Generic Sonar Model.

FOCUS simulator.

\section{Deep learning}

\begin{itemize}
\item AlphaGo - https://storage.googleapis.com/deepmind-media/alphago/AlphaGoNaturePaper.pdf
\item AlphaZero - Chess
\item Google’s Tacotron 2 text-to-speech system (autoregressive), based on WaveNet
\end{itemize}

Search spaces, learning speed,

Brain a good tutor - but AI should use operational definition and not a cognitive one. (Alan Turing / Tom M. Mitchell)

Autoregressive methods
Recurrent architectures are slower, and focus is on finding alternatives

ML by output classification
- Classification - discrete number of outputs (classes). Usually supervised.
- Regression - continuous outputs. Supervised.
- Clustering - unsupervised.
- Density estimation - finding the distribution of inputs in some space.
- Dimensionality reduction - simplifies inputs by mapping them to a lower-dimensional space. Topic modeling is related, finding human language similarities
- Learning to learn - learns its own inductive bias.
- Developmental learning / robot learning - generates its own sequences of learning situations to learn new skills through autonomous self-exploration and social interaction.

Keywords:
- Search space
- (deep) Reinforcement learning
  - Standard methods
    * Example: Deep Q-learning
  - Evolution strategies
    * Don't rely on gradients
    * Good when data isn't iid, error signals are sparse, and when need for exploration
    * Scale well with parallel hardware / machine clusters
- Supervised learning

Thinking fast and slow - Kahneman

Decision-based methods
- Supervised
  - Statistical methods
    - Parametric
      - Maximum likelihood
      - Bayesian estimation
    - Non-parametric
      - Window-method
      - Nearest-neighbor
      - Series expansion
  - Optimization methods
    - Linear discrimination functions
    - Generalized l.d.f.
- Non-supervised
  - Parametric methods
    - Maximum likelihood
    - Bayesian estimation
  - Cluster analysis
    - Optimization methods
    - Hierarchical methods

Structural methods
  - Syntactic methods
    - Syntax analysis
    - String matching

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\ifMonolithic\else\input{subfooter}\fi