### 2018-03-05 ThesisJo/preface.tex

\input{variables}\ifMonolithic\else\input{header}\input{subheader}\fi
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\chapter*{Preface}

This thesis has been submitted to the Faculty of Mathematics and Natural Sciences at the University of Oslo in partial fulfillment of the requirements for the degree \emph{Philosophiae Doctor} (Ph.D.). This work was carried out at the Digital Signal Processing and Image Analysis group at the Department of Informatics, University of Oslo, under the supervision of Professor Andreas Austeng and Professor Sverre Holm. The work was financed by the University of Oslo.

\section*{Acknowledgements}

\newpage
\chapter*{List of Publications}

\newpage
\chapter*{Related Publications}

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\ifMonolithic\else\input{subfooter}\fi