### 2018-10-11 ThesisJo/frontmatter.tex

%&header
% !TeX root = thesis
\input{subheader}
\endofdump

\ifRootBuild\else
  \input{glossary}
  \makeglossaries
\fi

\begin{document}

\begin{markdown}
Hey
\end{markdown}

\chapter*{Preface}

This thesis has been submitted to the Faculty of Mathematics and Natural Sciences at the University of Oslo in partial fulfillment of the requirements for the degree \emph{Philosophiae Doctor} (Ph.D.). This work was carried out at the Digital Signal Processing and Image Analysis group at the Department of Informatics, University of Oslo, under the supervision of Professor Andreas Austeng and Professor Sverre Holm. The work was financed by the University of Oslo.

\section*{Acknowledgements}

\newpage
\chapter*{Abstract}

\newpage
\chapter*{List of Publications}

This thesis is based on the following four journal papers, referred to by their roman literal:

\newcommand\pubEntry[4]{\item #2, "#1", \textit{#3}, #4}
{\renewcommand{\theenumi}{\textbf{\Roman{enumi}}}
\renewcommand\cleardoublepage{\newpage}
\begin{enumerate}
\pubEntry{%
An Optimized GPU implementation of the MVDR Beamformer for Active Sonar Imaging}%
{J.I. Buskenes, J.P. Åsen, C.-I.C. Nilsen and A. Austeng}%
{IEEE Journal of Oceanic Engineering}%
{Volume 40, Issue 2, April 2015}
%
\pubEntry{%
Implementing Capon Beamformer on a GPU for Real-Time Cardiac Ultrasound Imaging}%
{J.P. Åsen, J.I. Buskenes, C.-I.C. Nilsen, A. Austeng and S. Holm}%
{IEEE Transactions on Ultrasonics, Ferroelectrics, and Frequency Control}%
{Volume 61, Issue 1, January 2015}
%
\pubEntry%
{Low Complexity Adaptive Beamforming}%
{J.I. Buskenes, R.E. Hansen and A. Austeng}%
{IEEE Journal of Oceanic Engineering}%
{Volume 42, Issue 1, January 2017}
%
\pubEntry%
{A Real-Time Synthetic Aperture Sonar Simulator for Automatic Target Recognition: Notation and application}%
{J.I. Buskenes, H. Midelfart}%
{IEEE Journal of Oceanic Engineering}%
{Submitted DD.MM.2018}
\end{enumerate}
%}

%\newpage
\chapter*{Related Publications}

\begin{enumerate}
\pubEntry{%
A GPU sonar simulator for automatic target recognition}%
{J.I. Buskenes, H. Midelfart, Ø. Midtgaard}%
{Underwater acoustics international conference and exhibition}%
{June 2014}
%
\pubEntry{%
Adapting the minimum variance beamformer to a graphics processing unit for active sonar imaging systems}%
{J.I. Buskenes, J.P. Åsen, C.-I.C. Nilsen, and A. Austeng}%
{Journal of the Acoustical Society of America}%
{Volume 133, 2013}
%
\pubEntry%
{GPU-Based Adaptive Beamforming for Active Sonar Imaging}%
{J.I. Buskenes, J.P. Åsen, C.-I.C. Nilsen, and A. Austeng}%
{11th European Conference on Underwater Acoustics (ECUA)}%
{ISBN: 978-1-62276-192-0 p. 860-865, 2012}
%
\pubEntry%
{Implementing Capon Beamforming on the GPU for Real Time Cardiac Ultrasound Imaging}%
{J.P. Åsen, J.I. Buskenes, C.-I.C. Nilsen, A. Austeng, and S. Holm}%
{Proceedings - IEEE Ultrasonics Symposium}%
{p. 2133-2136, 2012}
%
\pubEntry%
{A Low Complexity Adaptive Beamformer for Active Sonar Imaging}%
{J.I. Buskenes, C.-I.C. Nilsen, and A. Austeng}%
{Underwater Acoustic Measurements: Technology and Results}%
{2011}
\end{enumerate}
}

\end{document}