### 2019-06-22 Simulator/article.tex
2019-06-21

\synctex=1
%%%%%%%%%%%%%%%%                                ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% CONDITIONALS %
%%%%%%%%%%%%%%%%                                ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\newif\ifPeerReview\PeerReviewfalse              % Whether to create the PeerReview version or
                                                % Journal version
\newif\ifOnlineColor\OnlineColortrue            % Compile online color version?

\newif\ifFlatArchive\FlatArchivefalse           % Whether archive is flat (messy) or contain 
                                                % subfolders for graphics etc.
\newif\ifFloatAtEnd\FloatAtEndfalse             % Available in PeerReview mode:
                                                % Place floats at end of document?
\newif\ifTODO\TODOtrue                        % Use todo notes?
\newif\ifOPTtodos\OPTtodosfalse
\newif\ifPhdDoc\PhdDocfalse

\newif\ifBuildBibliography\BuildBibliographytrue  % Whether to include the references.bib file or use inline references.
\newif\ifOverLeaf\OverLeaffalse                  % Using OverLeaf to edit?

\ifOnlineColor
   \newcommand\figPostfix{_online}              % Colored images online
\else
   \newcommand\figPostfix{_bw}                  % Black and white in journal
\fi

%%%%%%%%%%%%                                    ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% IEEEtran %
%%%%%%%%%%%%                                    ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\ifPeerReview
\documentclass[10pt,journal,draftclsnofoot,onecolumn]{IEEEtran}
% \newcommand\CLASSINPUTbaselinestretch{1.66}     % http://theoval.cmp.uea.ac.uk/~nlct/latex/thesis/node17.html
\else
\documentclass[journal]{IEEEtran}
\fi

% \RequirePackage[latin1]{inputenc}%              % Set input encoding (optionally latin1)
% \RequirePackage[T1]{fontenc}%                 % Set font encoding
% \usepackage[norsk]{babel}

% Single column review mode:
%\documentclass[12pt,journal,onecolumn]{IEEEtran}
%\newcommand\CLASSINPUTbaselinestretch{2}
%\RequirePackage{calc}
%\RequirePackage{fp} 
% Left margin = 1 inch + hoffset + oddsidemargin (or evensidemargin)
% Adding the 2cm gutter width to the odd/even side margins:
%\setlength\hoffset{0pt}
%\setlength\oddsidemargin{4cm}       % 4cm margins on the left side
%\addtolength\oddsidemargin{-1in}    % Subtract the initial 1 inch
%\setlength\textwidth{21cm-8cm}    % A4 width (21cm) minus margins on either side

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                  ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% IEEE ''APPROVED'' PACKAGES %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                  ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

%\usepackage[backend=biber,bibstyle=ieee]{biblatex}
%\addbibresource{references.bib}

\RequirePackage{amsfonts}
% \usepackage[makeroom]{cancel}
% \usepackage{cancel}
% \usepackage{ulem}
\usepackage{cite}

\ifCLASSINFOpdf
   \usepackage[dvips]{graphicx}                 % Might not work. Use 'latex' instead of 
   \ifFlatArchive\else                          % 'pdflatex'
      \graphicspath{./gfx/}
   \fi
\else
   \usepackage[dvips]{graphicx}
   \ifFlatArchive\else
      \graphicspath{./gfx/}
   \fi
\fi

\RequirePackage[table,dvipsnames,svgnames]{xcolor}

%\usepackage[cmex10]{amsmath}                    % cmex10 option to be IEEE explore compliant
\usepackage[cmex10]{mathtools}                    % cmex10 option to be IEEE explore compliant
\interdisplaylinepenalty=2500                   % Allows multiline equations to be broken
\RequirePackage{amssymb}

\RequirePackage{array}

\ifCLASSOPTIONcompsoc
   \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\else
   \usepackage[caption=false,font=footnotesize]{subfig}
\fi

% \usepackage{caption}
% \usepackage{subcaption}
\usepackage{color}
\usepackage{calc}
\usepackage{fp}

% 
% \ifCLASSOPTIONcaptionsoff                       % IEEE promoted hack to turn off captions from the 
%    \let\MYorigsubfloat\subfloat                 % subfloat package should the captionsoff option
%    \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}} % be specified.
% \fi

\ifFloatAtEnd
\ifCLASSOPTIONcaptionsoff                       % Places float at the end of the document when the
  \usepackage[nomarkers]{endfloat}              % captionsoff options is specified to IEEEtrans.cls
  \let\MYoriglatexcaption\caption               % (PeerReview mode)
  \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
\fi
\fi

% \usepackage{fixltx2e}                           % Fix some twocolumn float problems

% \usepackage{stfloats}                          % Allows: \begin{figure*}[!b]
                                                % (double column figures on top/bottom)

\usepackage{url}                                % Support for handling and breaking URLs

\usepackage{accents}
\usepackage{moresize}

% NOTE: PDF hyperlink and bookmark features are not required in IEEE
%       papers and their use requires extra complexity and work.
% \newcommand\MYhyperrefoptions{bookmarks=true,bookmarksnumbered=true,
% pdfpagemode={UseOutlines},plainpages=false,pdfpagelabels=true,
% colorlinks=true,linkcolor={black},citecolor={black},urlcolor={black},
% pdftitle={Optimal Window Design for the Low Complexity Adaptive Beamformer in Active Sonar Imaging},
% pdfsubject={},
% pdfauthor={Jo Inge Buskenes},
% pdfkeywords={adaptive beamforming, beamforming, complexity, sonar, active}}%
\ifCLASSINFOpdf
%    \usepackage[\MYhyperrefoptions,pdftex]{hyperref}
   \usepackage[pdftex,
    final,                %Keep hyperlink stuff even when in draft mode
    breaklinks=true,      %Break links when necessary
    linktocpage=true,     %Enable link to page?
    linkcolor=black,  %Colour of links to labels within document
    citecolor=black,      %Colour of links to the biliography
    filecolor=red,        %Colour of links to local files
    pagecolor=black,        %Colour of links to other pages withing document
    urlcolor=Blue,        %Colour of the links to external URLs
    colorlinks=true,      %??
    plainpages=false,     %Store roman/arabic numbering differently to avoid
    unicode=true,
    bookmarksopen=true,
%    bookmarksopenlevel=3,
    bookmarksnumbered=false ]{hyperref}
% 
% \else
%    \usepackage[\MYhyperrefoptions,breaklinks=true,dvips]{hyperref}%
\fi
%    \usepackage{breakurl}                        % Allows 'dvips' driver to break links

%%%%%%%%%%%%%%%%%%%%%%%                         ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% ADDITIONAL PACKAGES %       
%%%%%%%%%%%%%%%%%%%%%%%                         ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

% Exotic orchids
% https://visme.co/blog/wp-content/uploads/2016/01/63-1024x512.jpg
%\definecolor{cDarkBlue}{HTML}{192e5b}
%\definecolor{cBlue}{HTML}{1d65a6}
%\definecolor{cLightBlue}{HTML}{73a2c0}
%\definecolor{cGreen}{HTML}{00743f}
%\definecolor{cYellow}{HTML}{f2a104}
%\colorlet{cBase}{cLightBlue}
%\colorlet{cBaseMark}{cGreen}
%\colorlet{cPoint}{cYellow}
%\colorlet{cBasis}{cDarkBlue}

\ifOnlineColor
   \definecolor{tabBlue}{HTML}{AACCFF}
\else
   \definecolor{tabBlue}{HTML}{CCCCCC}
\fi

%\titleformat{\subsubsection}[hang]{\itshape}{\thesubsubsection}{0.6em}{}

%\makeatletter
%\xpatchcmd{\@sect}{%
%  \protected@edef\@svsec{\@seccntformat{#1}\relax}%
%}{%
%  \ifnum0=\pdfstrcmp{#1}{section}%
%  \def\@svsec{\circlednumber{\number\value{section}}\quad\relax}%
%  \else
%  \protected@edef\@svsec{\@seccntformat{#1}\relax}%
%  \fi
%  }{}{}
%\makeatother

%\titleformat{\subsubsection}{\Large\bfseries}{}{0em}{%
%\begin{picture}(0,0)
%\put(12,4){\circle{28}}
%\end{picture}\makebox[1.45em][c]{\thesubsubsection}\quad}

\usepackage{multirow}
\usepackage{tabularx}
\usepackage{collcell}
\usepackage{booktabs}
\usepackage{calc}
%\usepackage{ifthen}

%\usepackage[active,float]{preview}
%\PreviewEnvironment{gather} 
% \ifPeerReview
% \let\OldIncludegraphics{\includegraphics}
% \usepackage{letltxmacro}
% \LetLtxMacro{\OldIncludegrsaphics}{\includegraphics}
% \renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=\linewidth, #1]{#2}}
% \fi

% % \makeatletter
% % \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
% % \else\Gin@nat@width\fi}
% % \makeatother
% % \let\Oldincludegraphics\includegraphics
% % \renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

% \usepackage[maxfloats=40]{morefloats}
%\newcounter{todoidx}
%% \setcounter{todoidx}
%
%\ifTODO
%   \definecolor{todobackground}{rgb}{0.95,0.95,0.95}
%   \setlength\marginparsep{1pt}
%   \setlength\marginparwidth{35pt}
%   \newlength\marginparwidthsmall
%   \setlength\marginparwidthsmall{\marginparwidth}
%   \addtolength\marginparwidthsmall{-7pt}
%   \newcommand\todo[1]{%
%      \addtocounter{todoidx}{1}%
%      {\color{Red}\bf(\thetodoidx{})}%%\fbox{\bf\thetodoidx{}}}%
%      \marginpar{%
%         {\vspace*{-10pt}\color{Red}\fbox{\bf\thetodoidx{}}}\\%
%         \fcolorbox{red}{todobackground}{\parbox{\marginparwidthsmall}{\raggedright\scriptsize #1}}}}
%
%   \newcommand\todopar[1]{\fcolorbox{red}{white}{\parbox{0.97\linewidth}{#1}}}
%\else
%%    \usepackage[disable]{./todonotes} 
%   \newcommand\todo[1]{}
%\fi

\usepackage{listings}
%\usepackage[outputdir=build]{minted}

\lstset{
  %These 3 parameters should be re-specified for each listing:
   language=bash%
 , caption=%
 , label=%
% , extendedchars=false
%%  , inputencoding=utf8
%  %
%%  ,  fancyvrb=false % Beamer hates fancyvrb
%%   morecomment=[1]\ ,
%%   fvcmdparams=<\frac><2>,
% , backgroundcolor=\color{codeBackground}%
%  %
 , basicstyle=\small
% , keywordstyle=\color{DarkBlue}%
%% , identifierstyle=\color{green}%
% , commentstyle=\color{ForestGreen}\itshape%
%  %morecomment=[|]{//}
% , stringstyle=\color{Brown}%
%  %
%% , showspaces=false%
%% , showstringspaces=false%
%% , showtabs=false%
%  %
%% , frame=single                            %single, shadowbox, framed
 , frame=bt
%%  , xtopmargin=0pt%
%%  , xbottommargin=0pt%
%% , framesep=20pt
%% , rulesep=20pt
%% , lineskip=1ex
 , aboveskip=\baselineskip % space above listing
 , belowskip=\baselineskip % space below listing
 , framexleftmargin=1em%
% , framexrightmargin=0pt%
% , linewidth=\linewidth%
 , xleftmargin=1em%        %To vertically align the box with the rest of the text in the
% , xrightmargin=0em%       %document, rather than having the text aligned and the box overshoot.
% , tabsize=2%
% , abovecaptionskip=1\baselineskip
% , belowcaptionskip=0pt
%  %  
 , captionpos=b%     %Caption at bottom
%  % 
% , breaklines=true%  
% , breakatwhitespace=false%
%  %
 , mathescape=true%
%% , literate=*{-}{-}1
% , columns=fullflexible
%% , float
%% , extendedchars=false
%% , escapeinside={\$}{\$}
%%   escapebegin={|},
%%   escapeend={|},
%% , inputencoding=utf8x
%  %
%% , numbers=left%
%% , numberstyle=\tiny\color{SeaGreen}%
%% , stepnumber=1%
%% , numbersep=2em 
 }

\newcommand\gls[1]{}

\newenvironment{narrow}[2]{%
\begin{list}{}{%
\setlength{\topsep}{0pt}%
\setlength{\leftmargin}{#1}%
\setlength{\rightmargin}{#2}%
\setlength{\listparindent}{\parindent}%
\setlength{\itemindent}{\parindent}%
\setlength{\parsep}{\parskip}}%
\item[]}{\end{list}}

% \usepackage{float}

%% Cool blues (sea)
%\definecolor{deep_aqua}{HTML}{003b46}
%\definecolor{ocean}{HTML}{07575b}
%\definecolor{wave}{HTML}{66a5ad}
%\definecolor{sea_foam}{HTML}{c4dfe6}
%\colorlet{base}{wave}
%\colorlet{bright}{ocean}
%\colorlet{dark}{deep_aqua}
%
%% Invigorating
%% https://visme.co/blog/wp-content/uploads/2016/01/26-1024x512.jpg
%\definecolor{cDarkBlue}{HTML}{003b46}
%\definecolor{cBlue}{HTML}{07575b}
%\definecolor{cLightBlue}{HTML}{66a5ad}
%\definecolor{cYellow}{HTML}{c4dfe6}
%\definecolor{cRed}{HTML}{c4dfe6}
%\colorlet{base}{wave}
%\colorlet{bright}{ocean}
%\colorlet{dark}{deep_aqua}

%% Invigorating
%% https://visme.co/blog/wp-content/uploads/2016/01/26-1024x512.jpg
%\definecolor{cBlue}{HTML}{80add7}
%\definecolor{cGreen}{HTML}{0abda0}
%\definecolor{cWhite}{HTML}{ebf2ea}
%\definecolor{cYellow}{HTML}{d4dca9}
%\definecolor{cBrown}{HTML}{bf9d7a}
%\colorlet{cBase}{cBlue}
%\colorlet{cBaseMark}{cGreen}
%\colorlet{cPoint}{cYellow}
%\colorlet{cBasis}{cBrown}

%%%%%%%%%%                                      ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% MACROS %       
%%%%%%%%%%                                      ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\input{macros}

%\let\oldincludegraphics\includegraphics
%\renewcommand\includegraphics[2][]{%
%  \immediate\write18{./bin/laFigure #2 #1}%
%  \input{result}}%
%  
%% \DeclareMathOperator*{\argmin}{\text{arg}\;\text{min}}
%
%\newcommand\Fig[1]{Fig.~\ref{#1}}
%\newcommand\eq[1]{(\ref{#1})}
%
%\newcommand\Grey[1]{{\color{Grey}#1}}
%\newcommand\Red[1]{{\color{Red}#1}}
%\newcommand\Blue[1]{{\color{Blue}#1}}
%\newcommand\DarkBlue[1]{{\color{DarkBlue}#1}}
%\newcommand\LightBlue[1]{{\color{LightBlue}#1}}
%\newcommand\Brown[1]{{\color{Brown}#1}}
%\newcommand\Green[1]{{\color{Green}#1}}
%\newcommand\SeaGreen[1]{{\color{SeaGreen}#1}}
%\newcommand\Yellow[1]{{\color{yellow}#1}}
%\newcommand\Orange[1]{{\color{orange}#1}}
%
%\newcommand\nn{\nonumber\\}
%
%\newcommand\nmat[1]{\begin{matrix}#1\end{matrix}}
%\newcommand\bmat[1]{\begin{bmatrix}#1\end{bmatrix}}
%\newcommand\case[1]{\begin{cases}#1\end{cases}}
%\newcommand\textbox[2]{\footnotesize\text{\parbox{#1}{\centering\emph{#2}}}}
%
%\newcommand\rand{\text{rand}}
%\newcommand\randn{\text{randn}}
%\newcommand\rect{\text{rect}}
%\newcommand\sinc{\text{sinc}}
%\newcommand\tr{\text{tr}}
%\newcommand\adj{\text{adj}}
%
%% \newcommand\max{\text{max}}
%\newcommand\argmin[1]{\text{arg}\;\underset{#1}{\text{min}}}
%
%\newcommand\qqquad{\quad\qquad}
%\newcommand\qqqquad{\qquad\qquad}
%
%% \renewcommand\l[1]{\left#1}
%% \renewcommand\r[1]{\right#1}
%
%% {\text{\parbox{1.5cm}{\centering volume hyper- sphere}}}
%
%%Keyword colouring:
%\newcommand\kw[1]{#1}
%\newcommand\parm[1]{#1}%\color{Black}#1\color{Black}}
%
%\newcommand\of[1]{\scriptstyle(\parm{#1})\displaystyle}
%\newcommand\df[1]{\scriptstyle[\parm{#1}]\displaystyle}
%\newcommand\var[3]{#1_\text{#2}\of{#3}}
%
%\newcommand\diag{\text{diag}}
%
%% \raisebox{lift}[extend-above-baseline][extend-below-baseline]{text}
%\newcommand\mt[1]{\text{\emph{#1}}} %mt = mathtext
%\newcommand\mathnorm{\textstyle}
%\newcommand\mathbig[1]{\displaystyle#1\mathnorm}
%\newcommand\mathsmall[1]{\scriptstyle#1\mathnorm}
%\newcommand\mathtiny[1]{\scriptscriptstyle#1\mathnorm}
%\newcommand\sfrac[2]{\scriptstyle\raisebox{0.25pt}[0pt][0pt]{$\frac{#1}{#2}$}\mathnorm}
%\newcommand\nfrac[2]{\textstyle\frac{#1}{#2}\displaystyle}
%
%\newcommand\sumu[1]{\sum\limits^{#1}\;}
%\newcommand\suml[1]{\sum\limits_{#1}\;}
%\newcommand\sumb[2]{\sum\limits_{#1}^{#2}\;}
%
%\newcommand\produ[1]{\prod\limits^{#1}\;}
%\newcommand\prodl[1]{\prod\limits_{#1}\;}
%\newcommand\prodb[2]{\prod\limits_{#1}^{#2}\;}
%
%\newcommand\defeq{\overset{\underset{\mathrm{def}}{}}{=}}
%
%%Math macros:
%\newcommand\T{^{\scriptscriptstyle T}}
%\renewcommand\H{^{\scriptscriptstyle H}}
%
%\renewcommand\vec[1]{\boldsymbol{#1}}
%\newcommand\mat[1]{\boldsymbol{#1}}
%
%\newcommand\Om{O_\text{m}}
%\newcommand\Oa{O_\text{a}}
%\newcommand\Nl{N_\text{l}}
%\newcommand\Nk{N_\text{k}}
%\newcommand\1{\vec 1}
%\newcommand\I{\mat I}
%\renewcommand*\a{\vec a}
%\newcommand*\f{\vec f}
%\renewcommand*\i{\vec i}
%\renewcommand*\k{\vec k}
%\newcommand*\n{\vec n}
%\newcommand*\p{\vec p}
%\newcommand*\s{\vec s}
%\newcommand*\w{\vec w}
%\newcommand*\x{\vec x}
%\newcommand*\y{\vec y}
%
%\newcommand*\A{\mat A}
%\newcommand*\B{\mat B}
%% \newcommand*\C{\mat C}
%\newcommand*\E{\mat E}
%\renewcommand*\P{\mat P}
%\newcommand*\eP{\mat{\hat P}}
%\newcommand*\R{\mat R}
%\newcommand*\Ri{\R^{-1}}
%\newcommand*\eR{\mat{\hat R}}
%\newcommand*\eRi{\hat{\mat R}\;\!^{-1}}
%\newcommand*\Navg{N_\text{avg}}
%\newcommand*\W{\mat W}
%\newcommand*\X{\mat X}
%\newcommand*\Xd{\X_{\!\Delta}}
%\newcommand*\Y{\mat Y}
%
%\renewcommand\argmin{\text{argmin}}
%
%
%\renewcommand*\P{\mat P}
%\newcommand*\V{\mat V}
%\newcommand*\M{\mat M}
%
%\renewcommand*\L{\mat \Lambda}
%% \newcommand*\U{\mat U}
%% \renewcommand*\t{\mathtiny{^T}}
%% \newcommand*\h{\mathtiny{^H}}
%\renewcommand*\t{^T}
%\newcommand*\h{^H}
%
%\newcommand\minus{\scalebox{0.75}[1.0]{$-$}}
%
%% \usepackage{tikz}
%% \usetikzlibrary{shapes,snakes}
%\usepackage{amsmath,amssymb}
%% \usepackage{datatool}
%% \usepackage{glossaries}
%
%% \newenvironment{outline}
%% {\begin{itemize}}
%% {\end{itemize}}

\begin{document}

\input{document}

\end{document}