### 2019-06-28 ThesisJo/frontmatter.tex
2019-06-27

%&header
% !TeX root = thesis
\input{subheader}
\endofdump

\cleardoublepage

\vspace*{\fill}
\begin{center}
{\huge\bf Til Joakim} \\[1.5cm]
{\Large aldri gi opp}
\end{center}
\vspace*{\fill}

\clearpage

\ifRootBuild\else
  \input{glossary}
  \makeglossaries
\fi

\begin{document}
\pagestyle{normal}

\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}
% \cfoot{\thepage}

This thesis has been submitted to the Faculty of Mathematics and Natural Sciences at the University of Oslo in partial fulfillment of the requirements for the degree \emph{Philosophiae Doctor} (Ph.D.). This work was carried out at the Digital Signal Processing and Image Analysis group at the Department of Informatics, University of Oslo, under the supervision of Professor Andreas Austeng and Professor Sverre Holm. The work was financed by the University of Oslo.

\section*{Acknowledgements}

This journey seems to have started a different life ago. Along the way more appreciation has piled up in the depths of my heart than I believed possible. I can only hope to give a little back by extending my sincerest gratitude to

\emph{Professor Andreas Austeng}, for offering great advice, insights, ever-presence, persistence and unwaivering optimism---simply being the best supervisor imaginable;

\emph{Jon Petter Åsen}, for being a wonderful and joyful office mate, colleague, coauthor and friend during the first years;

\emph{Cherie}, my dear faraway friend, for the steady stream of encouraging thoughts, patience and help---asking nothing in return;

\emph{the Department of Informatics, IFE}, for initially granting me this opportunity and later extending it;

\emph{the Norwegian Defence Research Institute, FFI}, for providing data to study and being an amazing workplace;

\emph{Professor Roy Hansen} (FFI/IFE), for initially sparking my interest for sonars and this PhD;

\emph{Professor Sverre Holm} (FFI/IFE), for being my second supervisor, an inspiration, a bottomless pit of knowledge and an aid while I lectured signal processing;

\emph{Stig Synnes} (FFI/IFE), for getting me into volleyball and the mutual friendly nudges to finish writing;

\emph{Kenneth Gade} (FFI), for volleyball enthusiasm and the inspiration and help developing the notation in paper IV,

\emph{Carl-Inge C. Nielsen} (IFE), for the time, knowledge and flying start you offered when this all began;

\emph{Herman Midelfart (FFI)}, for getting a chance to work with you before your days were numbered;

\emph{other colleagues and students} at IFE and FFI, for making work and life enjoyable, every day;

\emph{Friskvernklinikken}, Asker, for showing me a path where I thought there were none;

\emph{Marthe Arnesen}, for the gesture you once made yourself, for what is and for what may come;

\emph{my family}, for all your support, love and presence---through good and bad;

(\emph{myself}, I could not have done it without you;)

\emph{Joakim}, for giving me purpose, perspective and joy. This is for you.

\newpage
\chapter*{Abstract}

\textit{Cognition:}\\The mental action or process or acquiring knowledge and understanding through thought, experience, and the senses.

\newpage
\chapter*{Included publications}
\addcontentsline{toc}{chapter}{Included publications}

This thesis is based on the following four journal papers, referred to by their roman literal:

\newcommand\pubEntry[4]{\item #2\\"#1"\\\textit{#3}\\#4}
{\renewcommand{\theenumi}{\textbf{\Roman{enumi}}}
\renewcommand\cleardoublepage{\newpage}
\begin{enumerate}
\pubEntry{%
An Optimized GPU implementation of the MVDR Beamformer for Active Sonar Imaging}%
{J.I. Buskenes, J.P. Åsen, C.-I.C. Nilsen and A. Austeng}%
{IEEE Journal of Oceanic Engineering}%
{Volume 40, Issue 2, April 2015}
%
\pubEntry{%
Implementing Capon Beamformer on a GPU for Real-Time Cardiac Ultrasound Imaging}%
{J.P. Åsen, J.I. Buskenes, C.-I.C. Nilsen, A. Austeng and S. Holm}%
{IEEE Transactions on Ultrasonics, Ferroelectrics, and Frequency Control}%
{Volume 61, Issue 1, January 2015}
%
\pubEntry%
{Low Complexity Adaptive Beamforming}%
{J.I. Buskenes, R.E. Hansen and A. Austeng}%
{IEEE Journal of Oceanic Engineering}%
{Volume 42, Issue 1, January 2017}
%
\pubEntry%
{Real-Time Synthetic Aperture Sonar Simulation for Automatic Target Recognition: Notation and Application}%
{J.I. Buskenes, H. Midelfart}%
{IEEE Journal of Oceanic Engineering}%
{Submitted DD.MM.2019}
\end{enumerate}
%}

%\newpage
\chapter*{Excluded Publications}
\addcontentsline{toc}{chapter}{Excluded publications}

\begin{enumerate}
\pubEntry{%
A GPU sonar simulator for automatic target recognition}%
{J.I. Buskenes, H. Midelfart, Ø. Midtgaard}%
{Underwater acoustics international conference and exhibition}%
{June 2014}
%
\pubEntry{%
Adapting the minimum variance beamformer to a graphics processing unit for active sonar imaging systems}%
{J.I. Buskenes, J.P. Åsen, C.-I.C. Nilsen, and A. Austeng}%
{Journal of the Acoustical Society of America}%
{Volume 133, 2013}
%
\pubEntry%
{GPU-Based Adaptive Beamforming for Active Sonar Imaging}%
{J.I. Buskenes, J.P. Åsen, C.-I.C. Nilsen, and A. Austeng}%
{11th European Conference on Underwater Acoustics (ECUA)}%
{ISBN: 978-1-62276-192-0 p. 860-865, 2012}
%
\pubEntry%
{Implementing Capon Beamforming on the GPU for Real Time Cardiac Ultrasound Imaging}%
{J.P. Åsen, J.I. Buskenes, C.-I.C. Nilsen, A. Austeng, and S. Holm}%
{Proceedings - IEEE Ultrasonics Symposium}%
{p. 2133-2136, 2012}
%
\pubEntry%
{A Low Complexity Adaptive Beamformer for Active Sonar Imaging}%
{J.I. Buskenes, C.-I.C. Nilsen, and A. Austeng}%
{Underwater Acoustic Measurements: Technology and Results}%
{2011}
\end{enumerate}
}

\end{document}