### 2019-10-03 Simulator/bio/herman_midelfart.tex
2019-10-03

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bio/herman.eps}}]{Herman Midelfart}
held a M.Sc. and Ph.D. degree (2003) in genomics from the Norwegian University of Science and Technology (NTNU), Norway. Since 2007 he advanced research on automatic target recognition (ATR) in synthetic aperture sonar (SAS) at the Norwegian Defence Research Establishment, Kjeller, Norway.

Herman passed away in 2015. This work was finalized in his memory.
\end{IEEEbiography}