### 2020-04-12 ThesisJo/frontmatter.tex

%&header
% !TeX root = thesis
\input{subheader}
\endofdump

\cleardoublepage

\vspace*{\fill}
\begin{center}
{\huge\bf Til Joakim} \\[1.5cm]
{\Large aldri gi opp}
\end{center}
\vspace*{\fill}

\clearpage

\ifRootBuild\else
  \input{glossary}
  \makeglossaries
\fi

\begin{document}
\pagestyle{normal}

\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}
% \cfoot{\thepage}

This thesis has been submitted to the Faculty of Mathematics and Natural Sciences at the University of Oslo in partial fulfillment of the requirements for the degree \emph{Philosophiae Doctor} (Ph.D.). This work was carried out at the Digital Signal Processing and Image Analysis group at the Department of Informatics, University of Oslo, under the supervision of Professor Andreas Austeng and Professor Sverre Holm. The work was financed by the University of Oslo.

\section*{Acknowledgements}

This journey seems to have started a different life ago. Along the way more appreciation has piled up in the depths of my heart than I believed possible. I can only hope to give a little back by extending my sincerest gratitude to the following:

\emph{Professor Andreas Austeng}, for offering great advice, insights, ever-presence, persistence and unwaivering optimism---simply being the best supervisor imaginable;

\emph{Jon Petter Åsen}, for being a wonderful and joyful office mate, colleague, coauthor and friend during the first years;

\emph{Cherie}, my dear faraway friend, for the steady stream of encouraging thoughts, patience and help---asking nothing in return;

\emph{The Department of Informatics, IFE}, for initially granting me this opportunity and later extending it;

\emph{The Norwegian Defence Research Institute, FFI}, for providing data to study and being an amazing workplace;

\emph{Professor Roy Hansen} (FFI/IFE), for initially sparking my interest for sonars and this PhD;

\emph{Professor Sverre Holm} (FFI/IFE), for being my second supervisor, an inspiration, a bottomless pit of knowledge and an aid while I lectured signal processing;

\emph{Stig Synnes} (FFI/IFE), for getting me into volleyball and the mutual friendly nudges to finish writing;

\emph{Kenneth Gade} (FFI), for volleyball enthusiasm and the inspiration and help developing the notation in paper IV,

\emph{Carl-Inge C. Nielsen} (IFE), for the time, knowledge and flying start you offered when this all began;

\emph{Herman Midelfart (FFI)}, for getting a chance to work with you before your days were numbered;

\emph{other colleagues and students} at IFE and FFI, for making work and life enjoyable, every day;

\emph{Friskvernklinikken}, Asker, for showing me a path where I thought there were none;

\emph{Marthe Arnesen}, for the gesture you once made yourself, for what is and for what may come;

\emph{My family}, for all your support, love and presence---through good and bad;

\emph{Myself}, I could not have done it without you ;)

\emph{Joakim}, for giving me purpose, perspective and joy. This is for you.

\newpage
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

\input{abstract}

\newpage
\chapter*{Included publications}
\addcontentsline{toc}{chapter}{Included publications}

This thesis is based on the following four journal papers, referred to by their roman literal:

\newcommand\pubEntry[4]{\item #2\\"#1"\\\textit{#3}\\#4}
{\renewcommand{\theenumi}{\textbf{\Roman{enumi}}}
\renewcommand\cleardoublepage{\newpage}
\begin{enumerate}
\pubEntry{%
An Optimized GPU implementation of the MVDR Beamformer for Active Sonar Imaging}%
{J.I. Buskenes, J.P. Åsen, C.-I.C. Nilsen and A. Austeng}%
{IEEE Journal of Oceanic Engineering}%
{Volume 40, Issue 2, April 2015}
%
\pubEntry{%
Implementing Capon Beamformer on a GPU for Real-Time Cardiac Ultrasound Imaging}%
{J.P. Åsen, J.I. Buskenes, C.-I.C. Nilsen, A. Austeng and S. Holm}%
{IEEE Transactions on Ultrasonics, Ferroelectrics, and Frequency Control}%
{Volume 61, Issue 1, January 2015}
%
\pubEntry%
{Low-Complexity Adaptive Sonar Imaging}%
{J.I. Buskenes, R.E. Hansen and A. Austeng}%
{IEEE Journal of Oceanic Engineering}%
{Volume 42, Issue 1, January 2017}
%
\pubEntry%
{Real-Time Synthetic Aperture Sonar Simulation for Automatic Target Recognition: Notation and Application}%
{J.I. Buskenes, H. Midelfart}%
{IEEE Journal of Oceanic Engineering}%
{Submitted DD.MM.2019}
\end{enumerate}
%}

%\newpage
\chapter*{Related Publications}
\addcontentsline{toc}{chapter}{Related publications}

\begin{enumerate}
\pubEntry{%
A GPU sonar simulator for automatic target recognition}%
{J.I. Buskenes, H. Midelfart, Ø. Midtgaard}%
{Underwater acoustics international conference and exhibition}%
{June 2014}
%
\pubEntry{%
Adapting the minimum variance beamformer to a graphics processing unit for active sonar imaging systems}%
{J.I. Buskenes, J.P. Åsen, C.-I.C. Nilsen, and A. Austeng}%
{Journal of the Acoustical Society of America}%
{Volume 133, 2013}
%
\pubEntry%
{GPU-Based Adaptive Beamforming for Active Sonar Imaging}%
{J.I. Buskenes, J.P. Åsen, C.-I.C. Nilsen, and A. Austeng}%
{11th European Conference on Underwater Acoustics (ECUA)}%
{ISBN: 978-1-62276-192-0 p. 860-865, 2012}
%
\pubEntry%
{Implementing Capon Beamforming on the GPU for Real Time Cardiac Ultrasound Imaging}%
{J.P. Åsen, J.I. Buskenes, C.-I.C. Nilsen, A. Austeng, and S. Holm}%
{Proceedings - IEEE Ultrasonics Symposium}%
{p. 2133-2136, 2012}
%
\pubEntry%
{A Low Complexity Adaptive Beamformer for Active Sonar Imaging}%
{J.I. Buskenes, C.-I.C. Nilsen, and A. Austeng}%
{Underwater Acoustic Measurements: Technology and Results}%
{2011}
\end{enumerate}
}

%\newpage
\chapter*{Symbol reference}
\addcontentsline{toc}{chapter}{Symbol reference}

%\begin{table}[H]\noindent%
%\begin{tabular}{>{$}l<{$} >{}l<{} l >{$}l<{$}}
\newcommand\pb[1]{\parbox[c]{9mm}{\textbf{\centering#1}}}
\newcommand\bb[1]{\mathbb{#1}}
%@{\hspace*{0\tabcolsep}}
\begin{longtable}{@{\hspace*{-\tabcolsep}}>{$}l<{$}   >{$\mathbb}l<{$}  l  >{$}l<{$}  @{\hspace*{-\tabcolsep}}>{\footnotesize}l }
\hspace*{-.5cm}\textbf{Symbol} & \textbf{Space} & \textbf{Description}                                              & \textbf{Equal to}                     \\\endhead
	\mathcal{A}          &                 & Name of geometrical object (e.g. an \textbf{A}rray)                     &                                       \\
	\ubar{A}             & \bb V^3\bb A    & Frame (configuration) of $\mathcal{A}$                                  & (\uvec{A},\udot{A})                   \\
	\uvec{A}             & \bb V^3         & Orientation of $\mathcal{A}$ (absolute, coordinate free)                &                                       \\
	\udot{A}             & \bb A           & Position of $\mathcal{A}$ (absolute, coordinate free)                   &                                       \\
	\hat{a}              & \bb V^3         & Steering vector (unit-length, coordinate-free)                          &                                       \\
	\dvec{a}             & \bb R^3         & Steering vector (unit-length, decomposed)                               &                                       \\
	\dvec{B}             & \bb C^{N_b,L}   & Butler/beam-space matrix (sparse DFT matrix)                            &                                       \\
	c                    & \bb R^+         & Wave propagation speed [m/s]                                            &                                       \\
	\Delta_m             & \bb R           & Delay applied to $m$th sensor (normalized by $f_s$)                     &                                       \\
	\dvec{\Delta}        & \bb R^M         & Delay vector, all sensors (normalized by $f_s$)                         &                                       \\
	f                    & \bb R           & Field strength in (\ref{eq_wave_equation})                              &                                       \\
	                     & \bb R^+         & Frequency (everywhere else) [Hz]                                        &                                       \\
   f_s                  & \bb R^+         & Temporal sampling frequency [Hz]                                        &                                       \\
	F                    & \bb R^+         & Temporal sampling frequency (normalized by $f_s$)                       & \frac{f}{f_s}                         \\
	\vec{k}              & \bb V^3         & Wave vector, wavenumber with direction                                  & \lVert k\rVert \hat{a} &(1)           \\
	\lVert\vec{k}\rVert  & \bb R^+         & Spatial angular frequency (wavenumber) [1/m]                            & \frac{2\pi}{\lambda},\frac{\omega}{c} \\
	K                    & \bb Z           & Number of temporal samples to average (one-sided)                       &                                       \\
	\lambda              & \bb R^+         & Wavelength [m]                                                          &                                       \\
	L                    & \bb N           & Number of elements in subarray                                          &                                       \\
	M                    & \bb N           & Number of sensors (hydrophones/microphones)                             &                                       \\
	n                    & \bb Z           & Temporal sample index                                                   & f_s t                                 \\
	N                    & \bb N           & Number of temporal samples                                              &                                       \\
	N_K                  & \bb N           & Number of temporal samples to average                                   & 2K+1                                  \\
	N_L                  & \bb N           & Number of subarrays                                                     & M-L+1                                 \\
	\vec{p}_{AB}         & \bb V^3         & Position vector (coordinate-free) from $\udot{A}$ to $\dot{B}$          & \udot{B}-\dot{A}                      \\
	\dvec{p}_{AB}^C      & \bb R^3         & Position vector from $\udot{A}$ to $\dot{B}$, represented in $\uvec{C}$ &                                       \\
	\phi                 & \bb R           & Azimuth steering angle                                                  &                                       \\
	\varphi              & \bb R           & Elevation angle                                                         &                                       \\
	\dvec{R}             & \bb C^{M,M}     & Spatial covariance matrix                                               &                                       \\
	\R_{\mathrm{BS}}     & \bb C^{M',M'}   & Spatial covariance matrix (beam-space)                                  &                                       \\
	\eR                  & \bb C^{L,L}     & Sample spatial covariance matrix                                        &                                       \\
	\eR_{\mathrm{BS}}    & \bb C^{N_b,N_b} & Sample spatial covariance matrix (beam-space)                           &                                       \\
	t                    & \bb R           & Time [s]                                                                & \frac{n}{f_s}                         \\
	T_s                  & \bb R           & Temporal sampling period [s]                                            & F_s^{-1}                              \\
	\theta               & \bb R           & Azimuth angle                                                           & \omega t - \vec k\vec p, \Omega n                    \\
	\omega               & \bb R^+         & Temporal angular frequency [1/s]                                        & 2\pi f                                \\
   \Omega               & \bb R^+         & Temporal angular frequency (normalized by $f_s$)                        & 2\pi F                                \\
	w_m                  & \bb C^N         & Weight on $m$th sensor                                                  &                                       \\
	\dvec{w}             & \bb C^{M,N}     & Weight set, all sensors                                                 &                                       \\
	x_m                  & \bb C^N         & Data (delayed, element-space) from $m$th sensor                         &                                       \\
	x_{\textrm{BS},m}    & \bb C^{N_b}     & Data (delayed, beam-space) from $m$th sensor                            &                                       \\
	\dvec{x}             & \bb C^{M,N}     & Data (delayed, element-space) from all sensors                          &                                       \\
	\dvec{x}_\mathrm{BS} & \bb C^{M,N_b}   & Data (delayed, beam-space) from all sensors                             &                                       \\
	z                    & \bb C^N         & Beamformer output                                                       &                                       \\
\multicolumn{4}{l}{{\footnotesize (1)}\small\textit\quad{}Assumes linear, homogeneous medium.}
\end{longtable}
%	y_m                  & \bb C^N         & Data (no delay, element-space) from $m$th sensor                        &                                       \\
%\end{tabular}
%\end{table}

%\newpage
\chapter*{Acronym reference}
\addcontentsline{toc}{chapter}{Acronym reference}

\begin{longtable}{l l}
	\textbf{Acronym} & \textbf{Description}                                                      \\
	ADC              & Analog-to-digital converter                                               \\
	ATR              & Automatic target recognition                                              \\
	AUV              & Autonomous underwater vehicle                                             \\
	BS-Capon         & Beam-space Capon (or MVDR, beamformer)                                    \\
	CPU              & Central processing unit                                                   \\
	CUDA             & Compute unified device architecture (from Nvidia)                         \\
	DAC              & Digital-to-analog converter                                               \\
	DAS              & Delay-and-sum (beamformer)                                                \\
	DFT              & Discrete Fourier transform                                                \\
	DMA              & Direct memory access                                                      \\
	ES-Capon         & Element-space Capon (or MVDR, beamformer)                                 \\
	FFT              & Fast Fourier transform                                                    \\
	FPGA             & Field-programmable gate array                                             \\
	FPR              & False positive rate                                                       \\
	FWHM             & Full width at half maximum                                                \\
	GPU              & Graphics processing unit                                                  \\
	HISAS1030        & High resolution synthetic aperture sonar, Kongsberg Maritime (model 1030) \\
	HPC              & High performance computing                                                \\
	HUGIN            & An autonomous underwater vehicle developed by Kongsberg Maritime          \\
	LCA              & Low complexity adaptive (beamformer)                                      \\
   MCM & Mine counter measures \\
	MKL              & Math kernel library (from Intel)                                          \\
	MVDR             & Minimum variance distortionless response (or Capon, beamformer)           \\
	NDC              & Normalized device coordinates                                             \\
	PCI (Express)    & Peripheral component interconnect (data bus)                              \\
	RAM              & Random access memory                                                      \\
	ROC              & Receiver operating curve                                                  \\
	SAS              & Synthetic aperture sonar                                                  \\
	SDFT             & Sliding discrete Fourier transform                                        \\
	SIGMAS+          & Name of an acoustic GPU-based simulator                                   \\
	SSS              & Sidescan sonar                                                            \\
	SM               & Streaming multi-processor \\
   TPR & True positive rate 
\end{longtable}

\end{document}